---
title: 'Implementing reproducibility in phonetic research: a computational workflow'
author: |
    | Stefano Coretta
    | *University of Manchester*
date: |
    | **mFiL 2017**
    | 28 April 2017
output: 
  beamer_presentation: 
    citation_package: natbib
    colortheme: spruce
    fonttheme: professionalfonts
    highlight: tango
    latex_engine: xelatex
    theme: Rochester
    fig_caption: false
bibliography: linguistics.bib
biblio-style: unified.bst
indent: true
header-includes:

- \frenchspacing
- \urlstyle{tt}
- \usepackage{graphicx}
- \setcitestyle{aysep={},notesep={:},citesep={,}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.cap="", fig.path = "figures/", fig.align = "center")
devtools::load_all("../speakr")
library(tidyverse)
theme_set(theme_light())
```

# Reproducible research

A piece of research is **reproducible** when, along with its *results*, the *data* and the *computational environment* that produced those results are made available to other researchers [@fomel2009].

# Reproducible research

\centering
\includegraphics[height = \textheight]{figures/reproducible.pdf}

<!--
# Reproducible research

![](figures/replicability.pdf)
-->

# Why should we care?

The \textcolor{red}{\textbf{problem}} [@sandve2013]:

- difficulty of reproduction
- difficulty of replication
- retracted papers (\url{http://retractionwatch.com})

The "Yokuts vowels" case [@weigel2002]:

- about **75%** of the data is contrived [@weigel2005, 149]
- some of the generalisations are **wrong** [@blevins2004a]

The \textcolor{green}{\textbf{solution}}:

- **Reproducible Research** (RR)

# Reproducible Research in linguistics

- **linked data** [@bird2003; @thieberger2004]

- **computational grammar** [@maxwell2005]

- RR in the Speech Sciences [@abari2012]
    - lack of scientific culture
    - inefficiency of infrastructure

# The workflow of phonetic research

- **Phase A**: scripting (Praat, @boersma2016)

- **Phase B**: results and analysis

- **Phase C**: dissemination

# Phase A: source code and documentation

Praat scripting:

- Atom editor (\url{https://atom.io})
    - syntax highlighting
    - autocompletion and snippets
- Literate Markdown 
    - tangle: `lmt` (\url{https://github.com/driusan/lmt})
    - weaving: `pandoc` (\url{http://pandoc.org})

# Atom

\centering
\includegraphics[width = \textwidth]{figures/atom.png}

# `lmt` (literate markdown tangler)

\centering
\includegraphics[width = \textwidth]{figures/lmt.png}

# `pandoc` (universal document converter)

\centering
\includegraphics[height = \textheight]{figures/pandoc.png}

# Phase B: the `speakr` package

`speakr` is an `R` [@r-core-team2015] package to aid Praat users (under development):

- aim: tangle and run Praat scripts from within `R`
- two main functions
    - `lmt()`: tangle a Praat script
    - `praatRun()`: run a Praat script

# Phase B: the `speakr` package

```{r, echo=TRUE, eval=FALSE}
# Tangle a Praat script
lmt("code/get-measurements.praat.md")

# Run the script
praatRun("code/get-measurements.praat")

# Read the results of the script
vowels <- read_csv("results/vowels.csv") %>%
    mutate_if(is.character, as.factor) %>%
    mutate(vowel = factor(vowel, c("i", "e", "a",
                                   "O", "u")))
```

# Phase B: the `speakr` package

```{r, include=FALSE}
lmt("code/get-measurements.praat.md")
praatRun("code/get-measurements.praat")
vowels <- read_csv("results/vowels.csv") %>%
    mutate_if(is.character, as.factor) %>%
    mutate(vowel = factor(vowel, c("i", "e", "a", "O", "u")))
```

```{r vowel-plot, out.width="\\textwidth"}
vowels %>%
    ggplot(aes(F2, F1, colour = vowel)) +
    geom_point() +
    scale_x_reverse(position = "top") +
    scale_y_reverse(position = "right") +
    stat_ellipse() +
    labs(title = "Vowel plot of one speaker of Italian", x = "F2 (Hertz)", y = "F1 (Hertz)")
```

# Phase C: dissemination

*There is no investigation without dissemination.*

\begin{flushright} Ricardo Bermúdez-Otero (p.c.) \end{flushright}

- `knitr` [@xie2014]
    - dynamic reports
    - reproducible documents

- GitHub (\url{https://github.com})
    - versioning system (`git`)
    - online repository
- Open Science Framework (\url{https://osf.io})
    - online repository (for data)

# Summary

- share data, source file(s), versioning

- increasing awareness of RR in linguistics

- Atom, `lmt`, `pandoc`, `speakr`, `knitr`

- this presentation (along with source code and data) is available at \url{https://github.com/stefanocoretta/reproducible-phonetics}

# Summary

\centering \Large
THANK YOU!

# References
